---
title: "Homework12"
author: "Nicole Phelan"
date: "2024-04-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Advanced ggplotting

```{r}
library(ggplot2)
library(dplyr)
library(patchwork)
```

In a grant proposal about hybridization between ferns Diplazium plantaginifolium and Diplazium ternatum, I want to include a plot of their overlapping elevation ranges. I want their ranges to be plotted on top of each other in different colors, and I want an intermediate color to display their shared elevation range.

```{r}
Species <- c("Diplazium plantaginifolium","Diplazium plantaginifolium","Diplazium ternatum","Diplazium ternatum")
El <- c(150,1600,500,2200) # lower and upper elevation bounds for each species
lab <- c("Species")
df <- data.frame(Species,El,lab)

p <- ggplot(data = df) +
  aes(x=lab,y=El,col=Species) +
  geom_segment(xend=lab, yend=1600, yend=2200,
               linewidth=15, 
               alpha=0.7)  +
  theme_classic(base_size = 20) +
  ylab("Elevation Range (m)") + 
  theme(axis.title.x=element_blank(),axis.text.x=element_blank())+
  scale_color_manual(values=c("darkgoldenrod1","darkturquoise"))
p
```

In a related project, I am assessing morphometric data for an undescribed herbarium specimen (Diplazium sp.) that may be a hybrid. Since hybrids tend to have intermediate morphology of their two parents, I am looking to make some plots that might demonstrate that.

```{r}
# Reading my data
sdata <- read.csv(file = "FernData.csv")

# Note: Due to a random update in patchwork, I am no longer able to display plots together by adding them

# Plotting the lamina length
p1 <- ggplot(data = subset(sdata, !is.na(LaminaLength.cm.))) +
  aes(x=Species, y=LaminaLength.cm., fill=Species) +
  geom_boxplot() +
  ylab("Lamina Length (cm)") +
  theme_classic(base_size = 15) +
  scale_fill_manual(values=c("coral2","goldenrod1","darkturquoise","lightpink")) +
  geom_point(position=position_jitter(width=0.1, height=0.7),
                                      color="black", alpha= 0.7, size=2) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

# Plotting the petiole length
p2 <- ggplot(data = subset(sdata, !is.na(Petiole.Length.cm.))) +
  aes(x=Species, y=Petiole.Length.cm., fill=Species) +
  geom_boxplot() +
  ylab("Petiole Length (cm)") +
  theme_classic(base_size = 15) +
  scale_fill_manual(values=c("plum","coral2","goldenrod1","darkturquoise","lightpink")) +
  geom_point(position=position_jitter(width=0.1, height=0.7),
                                      color="black", alpha= 0.7, size=2) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

# Plotting the lamina to petiole ratio
p3 <- ggplot(data = subset(sdata, !is.na(Petiole.Length.cm.) & !is.na(LaminaLength.cm.))) +
  aes(x=Species, y=LaminaLength.cm./Petiole.Length.cm., fill=Species) +
  geom_boxplot() +
  ylab("Lamina Length: Petiole Length") +
  theme_classic(base_size = 13) +
  scale_fill_manual(values=c("coral2","goldenrod1","darkturquoise","lightpink")) +
  geom_point(position=position_jitter(width=0.1, height=0.7),
                                      color="black", alpha= 0.7, size=2) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

p1
p2
p3

```

To measure leaf width I played around with 2 different ways to measure width (one was perpendicular and the other was along the branching veins). Let's see how different those graphs look from each other.

```{r}
# Measuring width along the leaf veins
p4 <- ggplot(data = subset(sdata, !is.na(LeafWidthAlong.cm.))) +
  aes(x=Species, y=LeafWidthAlong.cm., fill=Species) +
  geom_violin() +
  ylab("Lamina Width (cm)") +
  theme_classic(base_size = 15) +
  scale_fill_manual(values=c("plum","coral2","goldenrod1","darkturquoise","lightpink")) +
  geom_point(position=position_jitter(width=0.1, height=0.7),
                                      color="black", alpha= 0.7, size=2) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

# Measuring width perpendicular to the rachis
p5 <- ggplot(data = subset(sdata, !is.na(LeafWidthPerpendicular.cm.))) +
  aes(x=Species, y=LeafWidthPerpendicular.cm., fill=Species) +
  geom_violin() +
  ylab("Lamina Width (cm)") +
  theme_classic(base_size = 15) +
  scale_fill_manual(values=c("plum","coral2","goldenrod1","darkturquoise","lightpink")) +
  geom_point(position=position_jitter(width=0.1, height=0.7),
                                      color="black", alpha= 0.7, size=2) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

p4
p5
```

Measuring leaf width along the leaf veins definitely seems to separate out the species a little more than perpendicular width. Next, let's see how this might affect the length to width ratio.

```{r}
# Measuring width along the leaf veins
p6 <- ggplot(data = subset(sdata, !is.na(LL_WA))) +
  aes(x=Species, y=LL_WA, fill=Species) +
  geom_violin() +
  ylab("Lamina Length to Width Ratio") +
  theme_classic(base_size = 15) +
  scale_fill_manual(values=c("coral2","goldenrod1","darkturquoise","lightpink")) +
  geom_point(position=position_jitter(width=0.1, height=0.7),
                                      color="black", alpha= 0.7, size=2) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

# Measuring width perpendicular to the rachis
p7 <- ggplot(data = subset(sdata, !is.na(LL_WP))) +
  aes(x=Species, y=LL_WP, fill=Species) +
  geom_violin() +
  ylab("Lamina Length to Width Ratio") +
  theme_classic(base_size = 15) +
  scale_fill_manual(values=c("coral2","goldenrod1","darkturquoise","lightpink")) +
  geom_point(position=position_jitter(width=0.1, height=0.7),
                                      color="black", alpha= 0.7, size=2) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

p6
p7
```

Different measuring techniques drastically changes the L:W ratio for D. obscurum. However, based on both plots, Diplazium sp. has a much lower L:W ratio than any of the parent candidates I've gotten measurements for so far.

A lot of these plots did not look good as facet plots because there were too many groups. For the following facet plots, I am going to try and focus on just Diplazium sp., D. plantaginifolium, and D. urticifolium.

```{r}
cdata <- sdata[which(sdata$Species!="Diplazium cristatum" &sdata$Species!="Diplazium obscurum"),]

# Same data as p6
p8 <- ggplot(data = subset(cdata, !is.na(LL_WA))) +
  aes(x=Species, y=LL_WA, fill=Species) +
  geom_violin(show.legend = F) +
  ylab("Lamina Length to Width Ratio") +
  theme_classic() +
  scale_fill_manual(values=c("goldenrod1","darkturquoise","lightpink")) +
  geom_point(position=position_jitter(width=0.1, height=0.7),
                                      color="black", alpha= 0.7, size=2,show.legend = F) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

# Same data as p7
p9 <- ggplot(data = subset(cdata, !is.na(LL_WP))) +
  aes(x=Species, y=LL_WP, fill=Species) +
  geom_violin() +
  ylab("Lamina Length to Width Ratio") +
  theme_classic() +
  scale_fill_manual(values=c("goldenrod1","darkturquoise","lightpink")) +
  geom_point(position=position_jitter(width=0.1, height=0.7),
                                      color="black", alpha= 0.7, size=2) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

p8 + p9

```
