---
title: "Homework6"
author: "Nicole Phelan"
date: "2024-02-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Bombus Bumblebee Seasonal Variation

#### Finding the Best Fit Distribution
#### and Simulating a New Data Set

```{r, EVAL=FALSE}
library(ggplot2) 
library(MASS)
library(tidyverse)
```

#### Creating the Data Set
```{r}
# Importing the data set
z <- read.table(file="Burnham_field_data_bombus_seasonal_variation_Dataset.csv",
                header=TRUE,
                sep=",")

# Data cleaning NA values
z <- na.omit(z)
# Naming the variable of interest
z$myvar <- z$pathogen_load
# Converting the data to log function and eliminating excessive 0 values (that are skewing the data)
z <-z%>%filter(pathogen_load>0)%>%
  mutate(myvar=log(pathogen_load))

# Structure of the pathogen load variable
str(z$myvar)

# Summarizing stats of the data of interest
summary(z$myvar)
# How many 0s are in the original data (not the log version)
# This is what I removed through data cleaning
sum(z$pathogen_load>0)
```

#### Trying Out Different Distribution Fits

```{r}
# Plotting the histogram of the pathogen load
p1 <- ggplot(data=z, aes(x=myvar, y=after_stat(density))) +
  geom_histogram(color="grey60",fill="cornsilk",linewidth=0.2) 
print(p1)
# Adds geom density to smooth out the shape of the raw data
p1 <- p1 + geom_density(linetype="dotted",linewidth=0.75)
print(p1)
# Normal distribution fitted to the pathogen load data
normPars <- fitdistr(z$myvar,"normal")
# Shows the mean and sd being used for the fitted distribution
print(normPars) 
# Structure of normal distribution best fit
str(normPars)
# Estimated mean (truncated from initial print function)
normPars$estimate["mean"] # note structure of getting a named attribute

# Creating sd and mean variables for normal distribution function
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

# Generating possible data from the pathogen load data
xval <- seq(0,max(z$myvar),len=length(z$myvar))

# Creating and plotting the normal distribution function over the histogram of the pathogen load data
stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myvar), args = list(mean = meanML, sd = sdML))
p1 + stat

# Creating the variables for exponential probability density curve
expoPars <- fitdistr(z$myvar,"exponential")
rateML <- expoPars$estimate["rate"]
# Creating and plotting the exponential probability distribution function over the histogram of the pathogen load data and normal distribution function
stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myvar), args = list(rate=rateML))
p1 + stat + stat2

# Adding the uniform density curve to the mix
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myvar), args = list(min=min(z$myvar), max=max(z$myvar)))
p1 + stat + stat2 + stat3

# Finally the gamma density function
# Creating the variables first
gammaPars <- fitdistr(z$myvar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]
# Plotting the gamma function with everything else
stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myvar), args = list(shape=shapeML, rate=rateML))
p1 + stat + stat2 + stat3 + stat4

# A New plot with the beta probability density
# Cannot be added to original graph because pathogen load data needs to be re-scaled
# Code for the new histogram
pSpecial <- ggplot(data=z, aes(x=myvar/(max(myvar + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) + # the re-scaling step
  geom_density(linewidth=0.75,linetype="dotted")

#Now for the beta function
betaPars <- fitdistr(x=z$myvar/max(z$myvar + 0.1),start=list(shape1=1,shape2=2),"beta")
# Creating all our variables
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]
# Putting it all together
statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myvar), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

#### Simulating a New Data Set

```{r}
# Creating a Data Frame with Random Gamma Values
simData <- rgamma(n=length(z$myvar),shape=shapeML,rate=rateML)
id <- 1:length(z$myvar)
sim_frame <- data.frame(simData,id)

str(simData)
summary(simData)

# Plotting the Simulated Data with the Gamma Function 
p2 <- ggplot(data=sim_frame, aes(x=simData, y=after_stat(density))) +
  geom_histogram(color="grey60",fill="cornsilk",linewidth=0.2) 
p2 + stat4

# Plotting the Original Data with the Probability Density Curve
p1 + stat4
```

The gamma probability density curve does a pretty good job of modelling the original data, but it is not perfect. The mean value at which the distribution is centered is nearly consistent between the model and the real data (~13). However, the real data has a sharp decline in density at around 10, which might be a problem for representing real ecological constraints within the model.
