---
title: "Homework10"
author: "Nicole Phelan"
date: "2024-04-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## For Loops and Randomization Tests

#### Question 1 (As A Callable Function)

```{r}
##################################################################
# FUNCTION: Count_Zero
# packages: none
# count the number of 0s in a vector
# input: numeric_vector
# output: counter (number of 0s in the input vector)
# ----------------------------------------------------------------

Count_Zero <- function(numeric_vector){ 
  counter <- 0
  for (i in seq_along(numeric_vector)){
    if(numeric_vector[i]==0){
      counter = counter + 1
    }
  }
  return(counter) 
} 

random_vector <- c(0,4,0,4,5,6,7,5,7,0,7,0,6)
Count_Zero(numeric_vector=random_vector)
```

This can be written in less lines of code by getting rid of the Count_Zero function structure (ie. starting with counter <- 0), but I wanted to make this a function I could call later on.

#### Question 1 (Less Lines of Code)

```{r}
numeric_vector <- c(0,4,0,4,5,6,7,5,7,0,7,0,6)

counter <- 0
for (i in seq_along(numeric_vector)){
  if(numeric_vector[i]==0){
    counter = counter + 1
  }
}

print(counter)
```

#### Question 2 (As A Callable Function)

```{r}
##################################################################
# FUNCTION: Count_Zero1
# packages: none
# count the number of 0s in a vector
# input: numeric_vector
# output: counter (number of 0s in the input vector)
# ----------------------------------------------------------------

Count_Zero1 <- function(numeric_vector){ 
  counter <- 0
  counter <- length(numeric_vector[numeric_vector==0])
  return(counter) 
} 

random_vector <- c(0,4,5,6,7,5,7,0,7,0,6)
Count_Zero1(numeric_vector=random_vector)
```

This can be written in less lines of code by getting rid of the Count_Zero1 function structure (ie. starting with counter <- 0), but I wanted to make this a function I could call later on.

#### Question 2 (Less Lines of Code)

```{r}
numeric_vector <- c(0,4,5,6,7,5,7,0,7,0,6)

counter <- 0
counter <- length(numeric_vector[numeric_vector==0])

print(counter)
```

#### Question 3

```{r}
##################################################################
# FUNCTION: Matrix_Maker
# packages: none
# make matrix with the given dimensions where each output is row number * column number
# input: number of rows and number of columns
# output: matrix 
# ----------------------------------------------------------------

Matrix_Maker <- function(NRows,NCols){
  Vec <- rep(NA, NRows*NCols) # Make an vector of NAs the size of the matrix
  M <- matrix(data = Vec, ncol = NCols, nrow = NRows) # Turn vector into matrix
  M <- row(M)*col(M) # Add values
  return(M)
}

Matrix_Maker(NRows=4,NCols=5); Matrix_Maker(NRows=3,NCols=6)
```