---
title: "Homework10"
author: "Nicole Phelan"
date: "2024-04-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## For Loops and Randomization Tests

#### Question 1 (As A Callable Function)

```{r}
##################################################################
# FUNCTION: Count_Zero
# packages: none
# count the number of 0s in a vector
# input: numeric_vector
# output: counter (number of 0s in the input vector)
# ----------------------------------------------------------------

Count_Zero <- function(numeric_vector){ 
  counter <- 0
  for (i in seq_along(numeric_vector)){
    if(numeric_vector[i]==0){
      counter = counter + 1
    }
  }
  return(counter) 
} 

random_vector <- c(0,4,0,4,5,6,7,5,7,0,7,0,6)
Count_Zero(numeric_vector=random_vector)
```

This can be written in less lines of code by getting rid of the Count_Zero function structure (ie. starting with counter <- 0), but I wanted to make this a function I could call later on.

#### Question 1 (Less Lines of Code)

```{r}
numeric_vector <- c(0,4,0,4,5,6,7,5,7,0,7,0,6)

counter <- 0
for (i in seq_along(numeric_vector)){
  if(numeric_vector[i]==0){
    counter = counter + 1
  }
}

print(counter)

# As shown in class
counter <- 0
for (i in seq_along(numeric_vector)){
  counter <- counter + as.numeric(numeric_vector[i]==0)
}
```

#### Question 2 (As A Callable Function)

```{r}
##################################################################
# FUNCTION: Count_Zero1
# packages: none
# count the number of 0s in a vector
# input: numeric_vector
# output: counter (number of 0s in the input vector)
# ----------------------------------------------------------------

Count_Zero1 <- function(numeric_vector){ 
  counter <- 0
  counter <- length(numeric_vector[numeric_vector==0])
  return(counter) 
} 

random_vector <- c(0,4,5,6,7,5,7,0,7,0,6)
Count_Zero1(numeric_vector=random_vector)
```

This can be written in less lines of code by getting rid of the Count_Zero1 function structure (ie. starting with counter <- 0), but I wanted to make this a function I could call later on.

#### Question 2 (Less Lines of Code)

```{r}
numeric_vector <- c(0,4,5,6,7,5,7,0,7,0,6)

counter <- 0
counter <- length(numeric_vector[numeric_vector==0])

print(counter)

# As Shown in class
sum(numeric_vector==0)
```

#### Question 3

```{r}
##################################################################
# FUNCTION: Matrix_Maker
# packages: none
# make matrix with the given dimensions where each output is row number * column number
# input: number of rows and number of columns
# output: matrix 
# ----------------------------------------------------------------

Matrix_Maker <- function(NRows,NCols){
  Vec <- rep(NA, NRows*NCols) # Make an vector of NAs the size of the matrix
  M <- matrix(data = Vec, ncol = NCols, nrow = NRows) # Turn vector into matrix
  M <- row(M)*col(M) # Add values
  return(M)
}

Matrix_Maker(NRows=4,NCols=5); Matrix_Maker(NRows=3,NCols=6)
```

#### Question 4A

```{r}
##################################################################
# FUNCTION: Sim_Data
# packages: none
# simulate 3 groups of data with different means
# input: Mean1, Mean2, Mean3, Size (sample size of the groups)
# output: data frame of simulated data
# ----------------------------------------------------------------

Sim_Data <- function(Mean1,Mean2,Mean3,Size){
  Sim <- c(rnorm(n=Size,mean=Mean1),
           rnorm(n=Size,mean=Mean2),
           rnorm(n=Size,mean=Mean3))
  Groups <- c(rep("Group1",Size),rep("Group2",Size),rep("Group3",Size))
  SimData <- data.frame(Groups,Sim)
  return(SimData)
}

a <- Sim_Data(Mean1=50,Mean2=100,Mean3=30,Size=10)
print(a)

```

#### Question 4B

```{r}
##################################################################
# FUNCTION: Sim_Data_R
# packages: none
# simulate 3 groups with different means, reshuffle response variables, calculate new means
# input: Mean1, Mean2, Mean3 (sarting means), Size (sample size of the groups)
# output: vector with new means
# ----------------------------------------------------------------

Sim_Data_R <- function(Mean1,Mean2,Mean3,Size){
  Sim <- c(rnorm(n=Size,mean=Mean1),
           rnorm(n=Size,mean=Mean2),
           rnorm(n=Size,mean=Mean3))
  ReShuffled <- sample(x=Sim) # Reshuffles the Data
  Groups <- c(rep("Group1",Size),rep("Group2",Size),rep("Group3",Size))
  SimData <- data.frame(Groups,ReShuffled)
  G1 <- c(SimData[Groups=="Group1",]$ReShuffled)
  G2 <- c(SimData[Groups=="Group2",]$ReShuffled)
  G3 <- c(SimData[Groups=="Group3",]$ReShuffled)
  NewMeanVec <- c(Group1=mean(G1),Group2=mean(G2),Group3=mean(G3))
  return(NewMeanVec)
}

b <- Sim_Data_R (Mean1=50,Mean2=100,Mean3=30,Size=10)
print(b)
```

#### Question 4C

```{r}
##################################################################
# FUNCTION: Sim_Data_R_For
# packages: none
# simulate 3 groups with different means, reshuffle response variables, calculate new means, repeat 100 times
# input: Mean1, Mean2, Mean3 (sarting means), Size (sample size of the groups)
# output: data frame with new means and replicate number
# ----------------------------------------------------------------

Sim_Data_R_For <- function(Mean1,Mean2,Mean3,Size){
  DF <- data.frame()
  for (i in 1:100){
    Sim <- c(rnorm(n=Size,mean=Mean1),
             rnorm(n=Size,mean=Mean2),
             rnorm(n=Size,mean=Mean3))
    ReShuffled <- sample(x=Sim) # Reshuffles the Data
    Groups <- c(rep("Group1",Size),rep("Group2",Size),rep("Group3",Size))
    SimData <- data.frame(Groups,ReShuffled)
    G1 <- c(SimData[Groups=="Group1",]$ReShuffled)
    G2 <- c(SimData[Groups=="Group2",]$ReShuffled)
    G3 <- c(SimData[Groups=="Group3",]$ReShuffled)
    NewMeanVec <- c(Replicate=i,Group1=mean(G1),Group2=mean(G2),Group3=mean(G3))
    DF <-rbind(DF, NewMeanVec)
    
  }
   colnames(DF) <- c("Replicate", "NewMean1", "NewMean2", "NewMean3")
  return(DF)
}

c <- Sim_Data_R_For (Mean1=50,Mean2=100,Mean3=30,Size=10)
print(c)
```

#### Question 4D

```{r}

library(ggplot2)

p1 <- qplot(c$NewMean1, xlab = "New Means for Group 1", ylab = "Count")
p2 <- qplot(c$NewMean2, xlab = "New Means for Group 2", ylab = "Count")
p3 <- qplot(c$NewMean3, xlab = "New Means for Group 3", ylab = "Count")

print(p1);print(p2);print(p3)

```
