---
title: "Homework5"
author: "Nicole Phelan"
date: "2024-02-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1

```{r}

n_dims <- sample(3:10,size=1)
print(n_dims)

```

#### Question 1 Bullet Point 1

```{r}

con_int <- 1:(n_dims^2)
print(con_int)

```

#### Question 1 Bullet Point 2

```{r}

reshuffle <- sample(con_int)
print(reshuffle)

```

#### Question 1 Bullet Point 3

```{r}

sq_matrix <- matrix(reshuffle,nrow=n_dims)

```

#### Question 1 Bullet Point 4

```{r}

print(sq_matrix)

```

#### Question 1 Bullet Point 5

```{r}

transposed <- t(sq_matrix)

```

#### Question 1 Bullet Point 6

```{r}

print(transposed)

```

Transposing a matrix causes the rows to become columns and the columns to become rows.

#### Question 1 Bullet Point 7

```{r}

# Sum of the first row
first_row_sum <- sum(transposed[1,])
print(first_row_sum)

# Sum of the last row
last_row_sum <- sum(transposed[n_dims,])
print(last_row_sum)

# Mean of the first row
first_row_mean <- mean(transposed[1,])
print(first_row_mean)

# Mean of the last row
last_row_mean <- mean(transposed[n_dims,])
print(last_row_mean)

```

#### Question 1 Bullet Point 8

```{r}

eigen <- eigen(transposed)
print(eigen)

```

#### Question 1 Bullet Point 9

\$values refers to eigen values and $vectors refers to eigen vectors. The eigen vector is a type of vector that when multiplied by the eigen value, it results in a vector in the same or opposite direction as the original vector.

#### Question 1 Bullet Point 10

```{r}

typeof(eigen)

```

# Question 2

#### Question 2 Bullet Point 1

```{r}

my_matrix <- matrix(runif(16),nrow=4)
print(my_matrix)

```

#### Question 2 Bullet Point 2

```{r}

my_logical <- (runif(100)<0.5)
print(my_logical)

```

#### Question 2 Bullet Point 3

```{r}

my_letters <- sample(letters)
print(my_letters)

```

#### Question 2 Bullet Point 4

```{r}

newList <- list(my_matrix[2,2],my_logical[2],my_letters[2])
print(newList)

```

#### Question 2 Bullet Point 5

```{r}

typeof(newList[[1]]) # Determines data type of first list item
typeof(newList[[2]]) # Determines data type of second list item
typeof(newList[[3]]) # Determines data type of last list item
typeof(newList) # Determines data type of newList

```

#### Question 2 Bullet Point 6

```{r}

single_vec <- c(my_matrix[2,2],my_logical[2],my_letters[2])
print(single_vec)

```

#### Question 2 Bullet Point 7

```{r}

typeof(single_vec)

```

Everything in the list has been coerced into a character string.

# Question 3

#### Question 3 Bullet Point 1

```{r}

my_unis <- runif(26,0,10)
print(my_unis)

```

#### Question 3 Bullet Point 2

```{r}

my_letters <- sample(LETTERS)
print(my_letters)

```

#### Question 3 Bullet Point 3

```{r}

df <- data.frame(my_unis,my_letters)
print(df)

df[sample(nrow(df),4),1] <- NA
print(df)

```


#### Question 3 Bullet Point 4

```{r}

print(which(is.na(df)))

```

#### Question 3 Bullet Point 5

```{r}

library(dplyr)
alphabetized <- arrange(df, my_letters)
print(alphabetized)

```

#### Question 3 Bullet Point 6

```{r}

mean_first_col <- mean(alphabetized[,1],na.rm=TRUE)
print(mean_first_col)

```