---
title: "Homework 8"
author: "Nicole Phelan"
date: "2024-03-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Manipulation Using the `dplyr` Package

#### Question 1

```{r}
# Accessing useful packages
library(tidyverse)

# Accessing the iris data
data("iris")

#Examining the structure
glimpse(iris)
```

There are 5 variables (columns) for sepal length, sepal width, petal length, petal width, and species. There are 150 observations (rows) included in this data set.

#### Question 2

```{r}
# Creating a new data frame 
iris1 <- filter(iris, Species=="virginica" | Species=="versicolor", Sepal.Length>=6 & Sepal.Width>=2.5)
# Examining the structure
glimpse(iris1)
```

There are still 5 variables (columns) in the data set. However, after filtering out all others species than *I. versicolor* and *I. virginica* and creating criteria for sepal size, there are only 63 observations (rows).

#### Question 3

```{r}
# Getting rid of extra columns
iris2 <- select(iris1, Species, Sepal.Length, Sepal.Width)
# Examining the structure
glimpse(iris2)
```

There are still 63 observations (rows) like iris1, but now there are only 3 variables (columns).

#### Question 4

```{r}
# Largest to smallest sepal length
iris3 <- arrange(iris2, by=desc(Sepal.Length))
# Showing the first 6 rows
glimpse(iris3[1:6,])
```

#### Question 5

```{r}
# Adding column for sepal area
iris4 <- transmute(iris3, Species, Sepal.Length, Sepal.Width, Sepal.Area=Sepal.Length*Sepal.Width)
glimpse(iris4)
```

There are still 63 observations (rows) like iris2 and iris3, but now there are 4 variables (columns).

#### Question 6

```{r}
# Summarizing iris4
iris5 <- summarize(iris4, Average_Sepal_Length=mean(Sepal.Length),
                   Average_Sepal_Width=mean(Sepal.Width),
                   Sample_Size=nrow(iris4))
print(iris5)
```

#### Question 7

```{r}
# Summarizing iris4 for each species
iris_species <- group_by(iris4, Species)
iris6 <- summarize(iris_species, Average_Sepal_Length=mean(Sepal.Length),
                   Average_Sepal_Width=mean(Sepal.Width),
                   Sample_Size=n())
print(iris6)
```

#### Question 8

```{r}
# Reworking everything as a pipe statement
irisFinal <- iris %>%
  filter(Species=="virginica" | Species=="versicolor",
         Sepal.Length>=6 & Sepal.Width>=2.5) %>% # iris1
  select(Species, Sepal.Length, Sepal.Width) %>% # iris2
  arrange(by=desc(Sepal.Length)) %>% # iris3
  mutate(Sepal.Area=Sepal.Length*Sepal.Width) %>% #iris4
  group_by(Species) %>% # first step of iris6
  summarize(Average_Sepal_Length=mean(Sepal.Length),
                   Average_Sepal_Width=mean(Sepal.Width),
                   Sample_Size=n()) # second step of iris6
irisFinal
```

#### Question 9

```{r}
# Creating a longer data frame
irisLonger  <- iris %>%
  pivot_longer(cols=c(Sepal.Length, Sepal.Width, Petal.Length, Petal.Width), names_to="Measure", values_to="Value")
irisLonger
```