---
title: "Homework10"
author: "Nicole Phelan"
date: "2024-04-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## For Loops and Randomization Tests

#### Question 1

```{r}

##################################################################
# FUNCTION: Count_Zero
# packages: none
# count the number of 0s in a vector
# input: numeric_vector
# output: counter (number of 0s in the input vector)
# ----------------------------------------------------------------

Count_Zero <- function(numeric_vector){ # can remove this line
  counter <- 0
  for (i in seq_along(numeric_vector)){
    if(numeric_vector[i]==0){
      counter = counter + 1
    }
  }
  return(counter) # can remove this line
} # can remove this line

random_vector <- c(0,4,0,4,5,6,7,5,7,0,7,0,6)
Count_Zero(numeric_vector=random_vector)

```

This can be written in less lines of code by getting rid of the Count_Zero function structure (ie. starting with counter <- 0), but I wanted to make this a function I could call later on.

#### Question 2

```{r}

##################################################################
# FUNCTION: Count_Zero1
# packages: none
# count the number of 0s in a vector
# input: numeric_vector
# output: counter (number of 0s in the input vector)
# ----------------------------------------------------------------

Count_Zero1 <- function(numeric_vector){ # can remove this line
  counter <- 0
  counter <- length(numeric_vector[numeric_vector==0]) # this is the single line of code you need
  return(counter) # can remove this line
} # can remove this line

random_vector <- c(0,4,5,6,7,5,7,0,7,0,6)
Count_Zero1(numeric_vector=random_vector)
```

This can be written in less lines of code by getting rid of the Count_Zero1 function structure (ie. starting with counter <- 0), but I wanted to make this a function I could call later on.